Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2024-12-25 ms.qmd
Adjusting text in qmd too

Lastly, all models included a random slope of effect type (acute vs acclimation) to estimate the variance in  the magnitude of plastic changes (acute vs acclimation) across studies. Such an analysis is similar to analyses using an effect size that is a contrast between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ but is more powerful because it allows studies without acute responses to be included (see *Supplementary Materials*). With acclimation time and maximum temperature as moderators in all our models. Accounting for these in our meta-regression models, we proceeded to build separate models that tested our core questions. All estimates from our models are therefore conditioned on an average acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days) and an average maximum temperature (i.e., 23°C) across the dataset.

We first tested the extent to which acute and acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ effect sizes varied between habitat types (i.e., terrestrial, freshwater, and marine). Models included an interaction between effect type (i.e., acute or acclimation) and habitat (referred to as “Model 2”). Reduced mean $lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$ indicates that acclimation to thermal environments results in (partial) compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ indicates that organisms did not acclimate [@Havird2020-mj; @Seebacher2015-wi]. In contrast, a difference in $lnVR_{Q_{10_{acclimation}}}$ relative to $lnVR_{Q_{10_{acute}}}$ would show that changes in between-individual variation differ between acute responses and acclimation responses.

Second, we tested whether acute and acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ differed between whole-organism versus biochemical traits across habitats by fitting an model with an interaction between type, habitat and trait category (referred to as "Model 3").  We expected that whole-organism traits would be more likely to maintain variation in physiological function and be less likely to acclimate because whole-organism function relies on a greater number of biochemical reactions each with different thermal sensitivities [@fields2001protein; @angilletta2009thermal].

@@

#### **Acute and acclimation* for detailed trait categories across marine, freshwater and terrestrial taxa**

>>>>

Lastly, all models included a random slope of effect type (acute vs acclimation) to estimate the variance in  the magnitude of plastic changes (acute vs acclimation) across studies. Such an analysis is similar to analyses using an effect size that is a contrast between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ but is more powerful because it allows studies without acute responses to be included (see *Supplementary Materials*). With acclimation time and maximum temperature as moderators in all our models.

Accounting for these in our meta-regression models, we proceeded to build separate models that tested our core questions. All estimates from our models are therefore conditioned on an average acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days) and an average maximum temperature (i.e., 23°C) across the dataset.We first tested the extent to which acute and acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ effect sizes varied between habitat types (i.e., terrestrial, freshwater, and marine). Models included an interaction between effect type (i.e., acute or acclimation) and habitat (referred to as “Model 2”). Reduced mean $lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$ indicates that acclimation to thermal environments results in (partial) compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ indicates that organisms did not acclimate [@Havird2020-mj; @Seebacher2015-wi]. In contrast, a difference in $lnVR_{Q_{10_{acclimation}}}$ relative to $lnVR_{Q_{10_{acute}}}$ would show that changes in between-individual variation differ between acute responses and acclimation responses.

Second, we tested whether acute and acclimation $lnRR_{Q_{10}}$ and $lnVR_{Q_{10}}$ differed between whole-organism versus biochemical traits across habitats by fitting an model with an interaction between type, habitat and trait category (referred to as "Model 3"). A more detailed trait analysis is presented in the *Supplementary Materials*. We expected that whole-organism traits would be more likely to maintain variation in physiological function and be less likely to acclimate because whole-organism function relies on a greater number of biochemical reactions each with different thermal sensitivities [@fields2001protein; @angilletta2009thermal].

@@

#### **Acute and acclimation for detailed trait categories across marine, freshwater and terrestrial taxa**
